Real-space computation of $E$/$B$-mode maps I: Formalism, Compact Kernels, and Polarized Filaments

### 2018-07-23 articles/formalism/sections/discussion.tex
clean up and minor edits

Modified real-space operators with compact kernels could open several alternative analysis routes in the future.  No spin-harmonic transforms are necessary as the real space operators only rely on computing the Euler angles which can be done on the fly.  The radial functions (depending on $P_{\ell}^{2}$) need be tabulated only once at some determined resolution.  Especially given the Green's function interpretation of these operators, their implementation is trivially parallelizable over a compact domain, since the $E/B$ contribution from the Stokes charges in each pixel can be evaluated independently.  \revisit{Alternatively, the spatially-varying convolution kernel could be applied a polarized effective beam, as implemented in \cite{2011ApJS..193....5M} in a parallel scheme.}

The real space kernels could in principle be incorporated into the pointing matrices for map making, allowing maps of $E/B$ to be made directly from instrument data, without the need for Stokes parameter maps as intermediate products.
  The pointing matrix---projecting the maps into the time-ordered data at a point---requires the convolution version of the kernel, while the transpose pointing matrix---projecting the time-ordered data to the map---requires the Green's function kernel.
  The method would be similar to pixel-based strategies to deconvolve an instrument beam during map making \citep[e.g.][]{2010ApJS..187..212C}.  However, such a strategy for map making would force the pointing matrix to be much less sparse and it may not be practical.

>>>>

Modified real-space operators with compact kernels could open several alternative analysis routes in the future.  No spin-harmonic transforms are necessary as the real space operators only rely on computing the Euler angles which can be done on the fly.  The radial functions (depending on $P_{\ell}^{2}$) need be tabulated only once at some determined resolution.  Especially given the Green's function interpretation of these operators, their implementation is trivially parallelizable over a compact domain, since the $E/B$ contribution from the Stokes charges in each pixel can be evaluated independently.  Alternatively, the spatially-varying convolution kernel could be applied a polarized effective beam, as implemented in \cite{2011ApJS..193....5M} in a parallel scheme.

The real space kernels could in principle be incorporated into the pointing matrices for map making, allowing maps of $E/B$ to be made directly from instrument data, without the need for Stokes parameter maps as intermediate products.
  The pointing matrix---projecting the maps into the time-ordered data at a point---requires the convolution version of the kernel, while the transpose pointing matrix---projecting the time-ordered data to the map---requires the Green's function kernel.
  The method would be similar to pixel-based strategies to deconvolve an instrument beam during map making \citep[e.g.][]{2010ApJS..187..212C}. Such a strategy for map making would result in the pointing matrix in being much less sparse, hence making a practical implementation significantly challenging. Using the compact radial kernels might help with restoring some of the sparseness of the pointing matrix and this requires a more careful feasibility study.
  %However, such a strategy for map making would force the pointing matrix to be much less sparse and it may not be practical.

### 2018-07-23 articles/formalism/sections/real_space_formalism.tex
clean up and minor edits

\revisit{It turns out that the real space operations can \emph{also} be expressed in terms of the inverse rotation.  These two expressions lead to two sets of kernels that are conceptually different, although they both ultimately yield the same mathematical result.
The first set act like Green's functions, where, for example, a pixel of the Stokes parameters broadcasts or radiates an $E$/$B$ field.  The second set act like a convolving beams, gathering Stokes contributions to the $E/B$ fields at a point. On the equator, or in the flat sky approximation, these kernels are identical and this distinction is immaterial.  On the curved sky it matters, and the kernels are especially different near the poles.}

@@

The kernel $\mathcal{M}_G$ is the Green's function of the operator that transforms Stoke parameters to coordinate independent $E/B$.  The coordinate independence implies that the real and imaginary parts of the kernel do not vary with changes in the galactic latitude and longitude of the central pixel. In particular these functions are not distorted when a part of the domain overlaps with the poles, as can be seen in the first two rows of \fig{fig:vis_kernel}. From \eq{eq:qu2eb_radiation_concise}, we can see the $E/B$ patterns that Kronecker $\delta$-functions in the Stokes parameter pixels produce:
%
\beq
\bmat E= \textrm{Re}(\mathcal{M}_G) \\ B =\textrm{Im}(\mathcal{M}_G)  \emat  \leftarrow \bmat Q=\frac{\delta(\hat{n} - \hat{n}_q)}{\Delta \Omega}\\ U=0 \emat ;\qquad
\bmat E= -\textrm{Im}(\mathcal{M}_G) \\ B = + \textrm{Re}(\mathcal{M}_G)  \emat  \leftarrow \bmat Q=0 \\ U=\frac{\delta(\hat{n} - \hat{n}_q)}{\Delta \Omega} \emat.
\eeq
%
On the other hand, $\mathcal{M}^*_B$ is the Green's function of the operator that transforms $E/B$ to coordinate-dependent Stokes parameters $Q/U$.   The kernel $\mathcal{M}^*_B$ does not change with the central longitude, but varies as a function of galactic latitude. This latitude-dependent shape carries the  coordinate dependence of the Stokes parameters. From \eq{eq:eb2qu_radiation}, we $\mathcal{M}^*_B$ arises from $\delta$-functions in $E/B$:
%
\beq
\bmat Q= \textrm{Re}(\mathcal{M}^*_B) \\ U = \textrm{Im}(\mathcal{M}^*_B)  \emat  \leftarrow \bmat E=\frac{\delta(\hat{n} - \hat{n}_e)}{\Delta \Omega}\\ B=0 \emat;\qquad
\bmat Q= -\textrm{Im}(\mathcal{M}^*_B) \\ U = +\textrm{Re}(\mathcal{M}^*_B)  \emat  \leftarrow \bmat E=0 \\ B=\frac{\delta(\hat{n} - \hat{n}_e)}{\Delta \Omega} \emat\,.
\eeq
%
\revisit{Recall that the complex conjugates of these functions switch roles to form the convolution kernels. (For example, $\mathcal{M}_B$ is the convolution kernel for $Q/U \rightarrow E/B$.)}  It is also interesting to note that the radiation kernel ($\mm_{G}$) and the convolution kernel ($\mm_{B}$) become progressively identical as one approaches the equator as seen in  \fig{fig:vis_kernel}, which is a consequence of $\gamma \simeq -\alpha$ in the vicinity of the equator. The equator is the place on the sphere most akin to the flat sky case, where the radiation and convolution kernels are expected to be identical.

>>>>

It turns out that the real space operations can \emph{also} be expressed in terms of the inverse rotation.  These two expressions lead to two sets of kernels that are conceptually different, although they both ultimately yield the same mathematical result. The first set act like Green's functions, where, for example, a pixel of the Stokes parameters broadcasts or radiates an $E$/$B$ field.  The second set act like a convolving beams, gathering Stokes contributions to the $E/B$ fields at a point. On the equator, or in the flat sky approximation, these kernels are identical and this distinction is immaterial.  On the curved sky it matters, and the kernels are especially different near the poles.

@@

The kernel $\mathcal{M}_G$ is the Green's function of the operator that transforms Stoke parameters to coordinate independent $E/B$.  The coordinate independence implies that the real and imaginary parts of the kernel do not vary with changes in the galactic latitude and longitude of the central pixel. In particular these functions are not distorted when a part of the domain overlaps with the poles, as can be seen in the first two rows of \fig{fig:vis_kernel}. From \eq{eq:qu2eb_radiation_concise}, we can see the $E/B$ patterns that Kronecker $\delta$-functions in the Stokes parameter pixels produce:
%
\beq
\bmat E= \textrm{Re}(\mathcal{M}_G) \\ B =\textrm{Im}(\mathcal{M}_G)  \emat  \leftarrow \bmat Q=\frac{\delta(\hat{n} - \hat{n}_q)}{\Delta \Omega}\\ U=0 \emat ;\qquad
\bmat E= -\textrm{Im}(\mathcal{M}_G) \\ B = + \textrm{Re}(\mathcal{M}_G)  \emat  \leftarrow \bmat Q=0 \\ U=\frac{\delta(\hat{n} - \hat{n}_q)}{\Delta \Omega} \emat.
\eeq
%
On the other hand, $\mathcal{M}^*_B$ is the Green's function of the operator that transforms $E/B$ to coordinate-dependent Stokes parameters $Q/U$.   The kernel $\mathcal{M}^*_B$ does not change with the central longitude, but varies as a function of galactic latitude. This latitude-dependent shape carries the  coordinate dependence of the Stokes parameters. From \eq{eq:eb2qu_radiation}, we $\mathcal{M}^*_B$ arises from $\delta$-functions in $E/B$:
%
\beq
\bmat Q= \textrm{Re}(\mathcal{M}^*_B) \\ U = \textrm{Im}(\mathcal{M}^*_B)  \emat  \leftarrow \bmat E=\frac{\delta(\hat{n} - \hat{n}_e)}{\Delta \Omega}\\ B=0 \emat;\qquad
\bmat Q= -\textrm{Im}(\mathcal{M}^*_B) \\ U = +\textrm{Re}(\mathcal{M}^*_B)  \emat  \leftarrow \bmat E=0 \\ B=\frac{\delta(\hat{n} - \hat{n}_e)}{\Delta \Omega} \emat\,.
\eeq
%
Recall that the complex conjugates of these functions switch roles to form the convolution kernels (For example, $\mathcal{M}_G^*$ is the convolution kernel for $E/B \rightarrow Q/U$.).  It is also interesting to note that the radiation kernel ($\mm_{G}$) and the convolution kernel ($\mm_{B}$) become progressively identical as one approaches the equator as seen in  \fig{fig:vis_kernel}, which is a consequence of $\gamma \simeq -\alpha$ in the vicinity of the equator. The equator is the place on the sphere most akin to the flat sky case, where the radiation and convolution kernels are expected to be identical.
